Dependent Object Types

### 2012-09-22 doc/fool/article.tex
a bit more about expansion counterex

%% Relations
% Subtype 
\newcommand{\sub}{<:}
% Type assignment
\newcommand{\typ}{:}
% reduction
\newcommand{\reduces}{\;\rightarrow\;}
% well-formedness
\newcommand{\wf}{\;\mbox{\textbf{wf}}}
\newcommand{\wfe}{\;\mbox{\textbf{wfe}}}
\newcommand{\nswfe}{\mbox{\textbf{wfe}}}

@@

>>>>

%% Relations
% Subtype 
\newcommand{\sub}{<:}
% Type assignment
\newcommand{\typ}{:}
% reduction
\newcommand{\reduces}{\;\rightarrow\;}
% well-formedness
\newcommand{\wf}{\;\mbox{\textbf{wf}}}
\newcommand{\nswf}{\mbox{\textbf{wf}}}
\newcommand{\wfe}{\;\mbox{\textbf{wfe}}}
\newcommand{\nswfe}{\mbox{\textbf{wfe}}}

@@

First, let's illustrate why expansion does not always exist.

Here is the simplest such example:
\begin{align*}
&\mlnew z {\Top \mlrefine z {\\
&\gap \Ldecl {L} \Bot {z.L}\\
&}\ldefs{}}{}
\end{align*}
The type $z.L$ is $\nswf$ but not $\nswfe$. Indeed, there is no finite
derivation of an expansion for $z.L$, because by the
$\textsc{tsel-$\expand$}$ rule, in order to expand $z.L$, we need to
expand its upper bound, which is $z.L$! However, not that the object
creation expression for $z$ would not typecheck because subtyping is
regular wrt $\nswfe$, and so the type of a constructor in an object
creation expression must have $\nswfe$ type members, since we check
that the lower bound is a subtype of the upper bound for each type
member.

The example above relies on the $\textsc{tsel-wf$_2$}$, in order for
the upper bound to refer to the type member being declared. Here is an
example that does not rely on this rule, and that we will use below to
create a counterexample to preservation. Let
\begin{align*}
T_1 = \Top &\mlrefine z {&\\
&\gap\Ldecl A \Bot {z.B}&\\
&\gap\Ldecl B \Bot \Top&}\\
T_2 = \Top &\mlrefine z {&\\
&\gap\Ldecl A \Bot \Top&\\
&\gap\Ldecl B \Bot {z.A}&}
\end{align*}

Now, consider $T_1 \tand T_2$. The type is $\nswf$ and $\nswfe$, but
its members $A$ and $B$ are not $\nswfe$, because the expansion of
$T_1 \tand T_2$ with self-type $z$ has the set of declarations
$\{\Ldecl A \Bot {z.B},\;\Ldecl B \Bot {z.A}\}$ -- thus, to expand
$z.A$, we need to expand $z.B$, and to expand $z.B$, we need to expand
$z.A$! There is no finite derivation of expansions for the type
members $A$ and $B$.
